---
title: "P8105_HW2_jck2183"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

1. Assign a variable to a path of the dataset. 

```{r}
data_path <- "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```

Read the Mr. Trashwheel dataset.

```{r}
trashwheel_df <- read_xlsx(
    path = data_path, 
    sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)

median_2017 <- trashwheel_df %>% 
    filter(year == 2017) 

median(median_2017$sports_balls)
```

Read precipitation data! For 2018 and 2017.
```{r}
precip_2018 <-
	read_excel(
		data_path,
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year) 
    

sum(precip_2018$total)

precip_2017 <- 
	read_excel(
		path = data_path,
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

Combine annual precipitation.

```{r}
month_df <- 
	tibble(
		month = 1:12,
		month_name = month.name
	)

precip_df <- 
	bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```
Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2018? What was the median number of sports balls in a dumpster in 2017?

```{r}
# The trashwheel data frame contains 344 observations of 14 variables.The 2017 precipitation contains 12 observations of 3 variables and so does 2018 precipitation. The key variables can be the month or the year. In this practice, we used month as key variable. The total precipitation in 2018 is 70.33 while the median number of sports balls in a dumpster in 2017 is 8.
```


Problem 2

Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).

```{r}
prb2_path <- "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv"

prb2_df <- read_csv(prb2_path) %>% 
    janitor::clean_names() %>%
    select(c(line:ada), -c(exit_only, staffing, staff_hours)) %>% mutate(entry = recode(entry, 
                         `YES` = TRUE,
                         `NO` = FALSE))

dim(prb2_df)
```
Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

```{r}
# The dataset, after reading and cleaning,contains only 19 variables : line, station_name, station_latitude / longitude, 11 routes served, entry, vending, entrance type, and ADA compliance.The dimension of the resulting dataset includes 1868 rows x 19 columns. The data is tidy as the resulting dataset is arranged such that each variable is a column and each observation is a row.
```

Answer the following questions using these data:

1.How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St A/B/C/D; 125st 1; 125st 4/5); the distinct function may be useful here.
```{r}
prb2_1_df <- prb2_df %>% 
    distinct(station_name,line, .keep_all = TRUE) 
# There are 465 distinct rows in the resulting data frame, indicating that there are 465 distinct stations identified by both name and by line.  
```

2.How many stations are ADA compliant?

```{r}
prb2_1_df %>% 
    filter(ada = TRUE) %>% 
    count(ada)
#There are 84 stations are ADA compliant based on previous resulting data.
```


3.What proportion of station entrances / exits without vending allow entrance?

```{r}
prb2_path <- "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv"

prb2_3_df <- read_csv(prb2_path) %>% 
    janitor::clean_names() %>%
    mutate(entry = recode(entry, 
                         `YES` = TRUE,
                         `NO` = FALSE)) %>% 
    distinct(station_name,line, .keep_all = TRUE) %>% 
        filter(vending == "NO")

# The proportion of station entrances / exits without vending allow entrance is equal to 5/4 (?) = 1.25 or 5/2 =2.5 (?) 
```
Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

```{r}
ref_prb2_df <- prb2_df %>% 
    select(-c(station_latitude, station_longitude, entrance_type, entry, vending)) %>% 
    distinct(station_name,line, .keep_all = TRUE) %>% 
    gather(Route_Name, Route_Number, route1:route11) %>% 
    filter(`Route_Number` == "A" & `ada` == "TRUE")

#There are 17 distinct stations that serve the A train and are ADA compliant.
```

