---
title: "P8105_HW2_jck2183"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

1. Assign a variable to a path of the dataset. 

```{r}
data_path <- "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```

Read the Mr. Trashwheel dataset.

```{r}
trashwheel_df <- read_xlsx(
    path = data_path, 
    sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)

median_2017 <- trashwheel_df %>% 
    filter(year == 2017) 

median(median_2017$sports_balls)
```

Read precipitation data! For 2018 and 2017.
```{r}
precip_2018 <-
	read_excel(
		data_path,
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year) 
    

sum(precip_2018$total)

precip_2017 <- 
	read_excel(
		path = data_path,
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

Combine annual precipitation.

```{r}
month_df <- 
	tibble(
		month = 1:12,
		month_name = month.name
	)

precip_df <- 
	bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```
Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2018? What was the median number of sports balls in a dumpster in 2017?

```{r}
# The trashwheel data frame contains 344 observations of 14 variables.The 2017 precipitation contains 12 observations of 3 variables and so does 2018 precipitation. The key variables can be the month or the year. In this practice, we used month as key variable. The total precipitation in 2018 is 70.33 while the median number of sports balls in a dumpster in 2017 is 8.
```


## Problem 2

Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).

```{r}
prb2_path <- "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv"

prb2_df <- read_csv(prb2_path) %>% 
    janitor::clean_names() %>%
    select(c(line:ada), -c(exit_only, staffing, staff_hours)) %>% mutate(entry = recode(entry, 
                         `YES` = TRUE,
                         `NO` = FALSE))

dim(prb2_df)
```
Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

```{r}
# The dataset, after reading and cleaning,contains only 19 variables : line, station_name, station_latitude / longitude, 11 routes served, entry, vending, entrance type, and ADA compliance.The dimension of the resulting dataset includes 1868 rows x 19 columns. The data is tidy as the resulting dataset is arranged such that each variable is a column and each observation is a row.
```

Answer the following questions using these data:

1.How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St A/B/C/D; 125st 1; 125st 4/5); the distinct function may be useful here.
```{r}
prb2_1_df <- prb2_df %>% 
    distinct(station_name,line, .keep_all = TRUE) 
# There are 465 distinct rows in the resulting data frame, indicating that there are 465 distinct stations identified by both name and by line.  
```

2.How many stations are ADA compliant?

```{r}
prb2_1_df %>% 
    filter(ada = TRUE) %>% 
    count(ada)
#There are 84 stations are ADA compliant based on previous resulting data.
```


3.What proportion of station entrances / exits without vending allow entrance?

```{r}
nrow(filter(prb2_df,vending == "NO",entry == "TRUE"))/nrow(filter(prb2_df,vending == "NO"))

# The proportion of station entrances / exits without vending allow entrance is equal to 0.3770492.
```
Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

```{r}

A_station <- prb2_df %>% 
    select(-c(station_latitude, station_longitude, entrance_type, entry, vending)) %>% 
    distinct(station_name,line, .keep_all = TRUE) %>% 
    gather(Route_Name, Route_Number, route1:route11) %>% 
    filter(`Route_Number` == "A")

ref_prb2_df <- prb2_df %>% 
    select(-c(station_latitude, station_longitude, entrance_type, entry, vending)) %>% 
    distinct(station_name,line, .keep_all = TRUE) %>% 
    gather(Route_Name, Route_Number, route1:route11) %>% 
    filter(`Route_Number` == "A" & `ada` == "TRUE") 

#There are 17 distinct stations that serve the A train and are ADA compliant. There are 60 stations that serve A train. Of 60 stations, there are 17 distinct stations that are ADA compliant.
```

## Problem 3

First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

```{r}
pols_month_path <- "./data/pols-month.csv"
pols_month_df <- read_csv(pols_month_path) 
head(pols_month_df)

pols_month_df <- pols_month_df %>% 
    separate(mon, c("year", "month", "day"), sep = "-") %>% 
    mutate(month = as.numeric(month)) %>%
    mutate(month = month.name[month])  %>%
    mutate(president = rep_gop + rep_dem) %>% 
    select(-c("prez_dem", "prez_gop", "day"))
```
Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r}
snp_path <- "./data/snp.csv"
snp_df <- read_csv(snp_path) 
head(snp_df)


library(lubridate)
snp_df <- snp_df %>% 
    mutate(dt = as.Date(date, format = '%m/%d/%y')) %>%
    mutate(year = year(dt), month = month(dt), day = day(dt)) %>%
    select(c("year", "month", "close")) %>% 
    arrange(month, year) %>% 
    mutate(month = month.name[month]) %>% 
    mutate(year = as.character(year))

```

Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

```{r}
unemployment_path <- "./data/unemployment.csv"
unemployment_df <- read_csv(unemployment_path)   %>% 
    gather(month,Unemployment_Rate, Jan:Dec) %>% 
    mutate(month = recode(month, 
                         `Jan` = "January", `Feb` = "February", `Mar` = "March", `Apr` = "April", `May` = "May", `Jun` = "June", 
                         `Jul` = "July", `Aug` = "August", `Sep` = "September", `Oct` = "October", `Nov` = "November", `Dec` = "December")) %>% 
    mutate(Year = as.character(Year)) %>% 
    mutate(year = Year) %>% 
    select(-Year) %>% 
    relocate(year, month) %>% 
    arrange(month, year)
```
Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r}
m1_df <- left_join(pols_month_df, snp_df, by = c("year", "month")) 
m2_df <- merge(m1_df, unemployment_df, by = c("year", "month"))
```


Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

```{r}
#The first dataset combining the snp and pols_month data frames has 822 observations with 10 variables in total.This dataset contains all the 9 variables in pols and an extra variable "close value" from snp. Range of the years is 1947 to 2015. Names of the key variables are year and month. However, since the year range in snp is among 2019-2020, which failed to match with the key variables in pols_month data frame.

# The resulting dataset combines the first merged dataset with unemployment data frame. After merging with the key variables year and month, there are 11 variables and 810 observations. This dataset contains the previous 10 variables in the previous dataset while adding one more unemployment rate variable. The year range is 1948-2015.
```
Note: we could have used a date variable as a key instead of creating year and month keys; doing so would help with some kinds of plotting, and be a more accurate representation of the data. Date formats are tricky, though. For more information check out the lubridate package in the tidyverse.


